### 2016-02-11 ch05-initvals/initial-values.tex
Use make for figures, config compartments
Include initial values
Also quite a bit of pgfplots external trickery/hackery
to make sure the file compiles.
Some future work stuffies + labels
Add epigraphs
Include initvals review letter (prelim!)
Include review letter from initial values paper
Most of the review letter should go in the future.
Anyhow, at least the figures are already there, so we can add them
to the text.
Init-vals: more or less ready
Add TODOs for most of Johan's comments

\chapter{Initialization of parametric estimators}

>>>>

\chapter{Initialization of parametric estimators}
\label{sec:initvals}
\myEpigraph{Everything starts somewhere, though many physicists disagree. But people have always been dimly aware of the problem with the start of things. They wonder how the snowplough driver gets to work, or how the makers of dictionaries look up the spelling of words.}{Terry Pratchett}{Hogfather}
\gdef\thisDir{ch05-initvals}

\section{Introduction}\label{se:introd}

The task of interpreting measurement data from dynamic systems often involves the estimation of a transfer function (TF). 
Typical measurement techniques and identification strategies of transfer functions (TFs) or \glspl{FRF} are discussed in \citep{Schoukens1998,Schoukens2006LPM,Guillaume1996,Broersen1995,Pintelon2010LPM1,Antoni2007FRF}.
Application of these methods to real devices and systems is well-known in published literature, including \citep{Lim2010,Robinson1990,Behjat2010}.

Parametric identification of \gls{LTI} systems from input/output data has been well developed as evidenced by published literature~\citep{Pintelon2012,Ljung1999,Schoukens1999,Pintelon1998}. 
 Maximizing the likelihood function of the parameters probably yields the most popular, consistent and efficient estimator. 
 However, in a noisy environment, the \gls{MLE} engenders a nonlinear optimization which, at the very least, requires good initial estimates to avoid local optima.
 Such local optima have a detrimental effect on the quality of the estimated model.

 In the perspective of developing identification tools that require as little user interaction as possible, it is important to increase the probability of the optimization algorithm to produce good estimates, or preferably even the global optimum.
 However, when noisy input-output measurements are to be processed, the presence of noise obstructs a clear view of the actual system behavior, especially for poor signal-to-noise ratios (SNRs).
 In non-parametric identification, the presence of noise is typically mitigated by averaging multiple measurements or using smoothing techniques.

The main aim and contribution of this chapter is to demonstrate that smoothing a measured \gls{FRF} helps to avoid local optima during the parametric estimation of the \gls{MLE} of a transfer function, using deterministic optimization algorithms.
Hence, such techniques make it possible to increase the ``success rate'' (i.e. the probability that a good model, or even the global optimum, is obtained) of such a system identification step considerably.
%Such techniques, hence allow to increase the ``success rate'' (i.e. the probability that a good model, or even the global optimum, is obtained) of such a system identification step considerably.

In particular, two different smoothing techniques --- the time-truncated \gls{LPM}~\citep{Lumori2014TIM} and the \gls{RFIR}~\citep{Pillonetto2010,Chen2012} --- are tested for different \glspl{SNR} and different measurement record lengths of the input/output data of a few \gls{SISO} \gls{LTI} systems.
These are compared with the existing initialization schemes, namely: (i) the \gls{GTLS}, and (ii) the \gls{BTLS}.
%The existing initialization schemes compared with are the generalized total least squares (GTLS) and bootstrapped total least squares (BTLS).

The rest of the chapter is structured as follows. Section~\ref{se:ProbForm} covers the problem formulation. Section~\ref{se:MethodEg} discusses the methodology for obtaining the initial estimates, and their influence on the success rate of the \gls{MLE}. 
This is followed by Sections~\ref{se:Demo} and \ref{se:ExpMeas}, which are demonstrations of the improved initial estimates on simulation and experimental data, respectively.
In Section~\ref{se:Generality} the generality of the results is discussed briefly.
Ultimately, concluding remarks are presented in Section~\ref{se:Conclusion}.

% !TEX root =  InitialValuesPaper.tex

\section{Problem formulation}\label{se:ProbForm}
%\begin{itemize}
%\item
%Block schematic of the output error framework

%\JL{In this section, the assumptions on the system and the noise are given, and the associated Maximum Likelihood estimator, formulated in the frequency domain. The latter turns out to be a non-convex function in the parameters, requiring initial estimates. The obtention of those is discussed in the subsequent section.}

In this section, the assumptions on the system and the noise are presented, together with the associated MLE (formulated in the frequency domain), which turns out to be a non-convex function in the parameters, requiring initial estimates. The procedure for obtaining the initial estimates is discussed in the subsequent sections.

\subsection{System Setup Framework and Model}
\figref{fig:oesetup} depicts a schematic of the output error framework for  a generalized (single- or multiple-order) resonating, dynamic LTI discrete-time SISO system, subjected to a known white random noise input. The following mathematical derivations pertain to discrete time systems, but they can be extended to continuous time systems as well. The full mathematical model of the system is
\begin{equation}\label{lpmtd1}
y(t)=G_0(q^{-1})u_0(t)+H_0(q^{-1})e(t)
\end{equation}
%\comment{We don't have any results with colored noise yet!}
where $G_0(q^{-1})$ represents the dynamics of the system to be estimated, $u_0(t)$ is the input signal, $v(t)= H_0(q^{-1})e(t)$ is the noise source at the output, $H_0(q^{-1})$ is the noise dynamics, 
$e(t)$ is white Gaussian noise, and $q^{-m}$ is the backwards shift operator ($q^{-m}x(t)$ = $x(t-mT_{\mathrm{s}})$  with $m$ a positive integer and $T_{\mathrm{s}}$ the sampling time).
For white noise, $H_0 = 1$ and $e(t) = v(t)$.
Note that for simplicity, we only treat the discrete-time case (i.e. $t = n \cdot T_{\mathrm{s}}$ for integer values of $n$) theoretically in this chapter.
However, the generalization to continuous time is straightforward~\citep[Chapter 6]{Pintelon2012} and has been demonstrated in Section~\ref{se:ExpMeas}.

\begin{figure}[tbh] %top bottom here
\centering
\input{\thisDir/fig/oesetup.tikz}
\caption[Output-error set-up.]{SISO LTI discrete-time system in an output error setup.}
\label{fig:oesetup}
\end{figure}

Numerous parametric identification techniques are devoted to the development of parametric plants $G(q^{-1},\theta)$ and parametric noise models  $H(q^{-1},\theta)$, where  $\theta$ is the model parameters vector  \citep{Ljung1999,Soderstrom1989}.

In this chapter we assume that the output signal is disturbed by random noise, resulting in noisy error-prone data.
For the simulations, white noise is used.
It is also possible to apply the estimation procedure in this chapter to a system that is disturbed by colored noise, as will be demonstrated in \secref{se:ExpMeas}.

\subsection{Parametric System Model}

With reference to equation \eqref{lpmtd1}, the relation between the noiseless input and the output signals ($v(t)= 0$) is assumed to be of the form
\begin{equation}
    A(q^{-1}) y_0(t) = B(q^{-1}) u_0(t) \quad \forall t \in \left( - \infty, + \infty \right)
\label{ABpolys}
\end{equation}
where $A$ and $B$ are polynomials in $q^{-1}$. Thus, it follows from equation \eqref{lpmtd1} that
\begin{equation}
    G_0(q^{-1})= \frac{B(q^{-1})}{A(q^{-1})}
    \text{.}
\end{equation}

From~\citep[Section 6.3.2.]{Pintelon2012}, in the frequency domain, and for the DFTs (Discrete Fourier Transforms) of the windowed signals, equation \eqref{ABpolys} is of the form
\begin{align}
A(e^{-j\omega_k})Y_0(k) = B(e^{-j\omega_k})U_0(k) + I(e^{-j\omega_k})
\label{DFTspectra}
\end{align}
with $q^{-1}$ sampled in  $q_k^{-1} = e^{-j\omega_k}$, where $\omega_k = \frac{2\pi k}{N}$ are the DFT frequencies, and $I$ is a polynomial of order $\max(N_a,N_b) - 1$, which depends on the initial and end conditions. 
\TODO{explain init/end conditions in this context}
\begin{assumption}
It is assumed that the orders of the polynomials $A$, $B$ and $I$ are known.
\end{assumption}

\begin{remark}
The knowledge of the true model orders makes the analysis of the smoothers easier to carry out as we thus avoid extra calculation time required to perform model selection.
However, these smoothers can be used in conjunction with model order selection procedures.
\end{remark}

\subsection{Parametric Identification Algorithm}\label{se:paramIdentAlgo}

The maximum likelihood estimate of the parameter vector $\theta$  containing the coefficients of the $A$, $B$ and $I$  polynomials is obtained by solving the optimization problem
\begin{equation}
  \hat{\theta} = \arg\min_\theta V(\theta) \text{.}
\end{equation}
Since the noise $v(t)$ is assumed to be Gaussian, $V(\theta)$ accords to the weighted least squares cost function~\citep[Section 9.11]{Pintelon2012}:
\begin{equation}\label{eq:MLEcf}
V(\theta) = \sum_{k=1}^{N/2-1}\frac{|\varepsilon(k,\theta)|^2}{\sigma_\varepsilon^2(k,\theta)}
\end{equation}
for $N$ measurements and when $\varepsilon$ denotes the error in equation \eqref{DFTspectra},  \emph{viz}.:
\begin{align}
\varepsilon(k,\theta) = A(k,\theta)Y(k) - B(k,\theta)U(k) - I(k,\theta)\text{.}
\end{align}
The variance of the error is of the form:
\begin{equation}\label{eq:sigmaEps}
\begin{split}
\sigma_\varepsilon^2(k,\theta) 
  &=  |A(k,\theta)|^2\sigma_Y^2(k) 
   +  |B(k,\theta)|^2\sigma_U^2(k) \\
  &- 2\;\mathrm{real} \left( A(k,\theta) \overline{B(k,\theta)} \sigma_{YU}(k) \right)
\end{split}
\end{equation}
where $\sigma_Y^2$ is independent of $k$ for white noise, %assumed to be a known error variance \comment{JL: why should the error variance be known? where do you use that? how do you compute it?} of the measured output spectrum $Y$ \comment{JL: relate the output error variance with the variance of $v(t)$. Isn't it the same thing?},
%\comment{this is true for white noise, and for the scaling of the DFT by $\sqrt{N}$. I think it is more clear to use the frequency domain variance instead: $\sigma_Y^2$}
and the input error variance is assumed to be zero, i.e. $\sigma^2_U=0$, and similarly, the covariance $\sigma_{YU} = 0$.
Here, $\overline{B}$ denotes the complex conjugate of $B$.
%\comment{use $\sigma^2_U$ instead}

Consequently, $V(\theta)$ is a non-quadratic function of $\theta$ which, in general, results in a non-convex optimization problem. 
The Levenberg-Marquardt~\citep{Marquardt1963} algorithm (see Algorithm~\ref{LMalgo}) is used to solve this optimization problem deterministically.
Such an approach requires good initial estimates of $\theta$ to avoid inherent local optima, which is the focus of this chapter.

\begin{algorithm}
\caption{Levenberg-Marquardt~\citep{Marquardt1963}, \citep[Sec. 9.L.4]{Pintelon2012}}\label{LMalgo}
\begin{algorithmic}[1]
  \Statex Cost function \eqref{eq:MLEcf} is rewritten as $V(\theta) = \epsilon^{\TT} \epsilon$. 
  \Statex The iterations start at $\atIter[0]{\theta} = \theta^{\mathrm{init}}$ and $\lambda = \max(\sigma)/100$.
%\Function{LevenbergMarquardt}{$V(\theta), \atIter[0]{\theta}$}
   \For{$i$ \textbf{in} $1 \to \infty$}
      \State $U \, \Sigma \, W^{\TT} \gets \mathtt{svd}\left( \frac{\partial \epsilon}{\partial \theta}\right)$ with $\mathrm{diag}(\Sigma) = [\sigma_1, \ldots, \sigma_n]$
      % \State \textbf{if} $i=1$ \textbf{ then} $\lambda \gets \max(\sigma) / 100$
      \State $\sigma_k \gets \sigma_k / (\sigma_k^2 + \lambda^2) \quad \forall k \in \{1, \ldots, n\}$
      \State $\delta\theta \gets - W  \Sigma^{\TT}  U^{\TT} \epsilon$
      \State $\atIter[i]{\theta} \gets (\atIter[i-1]{\theta} + \delta\theta) / \norm{\atIter[i-1]{\theta} + \delta\theta} $
      \If{$V(\atIter[i]{\theta}) > V(\atIter[i-1]{\theta})$}
          \State $\lambda \gets 10 \lambda$ and then retry at $i \gets i - 1$
      \Else
          \State $\lambda \gets 0.4 \lambda$
      \EndIf
      \State \textbf{if} $\norm{\delta\theta}/\norm{\atIter[i]{\theta}} < n_{\theta}  \cdot 10^{-6} $ \textbf{then} \textbf{return} $\hat{\theta} \gets \atIter[i]{\theta}$
   \EndFor
%\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{remark}
Alternatively, stochastic optimization algorithms~\citep{Spall2012,Press2007} are less likely to get stuck in local optima since many randomized initial estimates are tried.
This obviously comes with a huge computational cost compared to `classical' deterministic schemes where only a single or a few initial estimates are used to start the expensive iterative optimization procedure.
Stochastic optimization is not discussed in the remainder of this chapter.
\end{remark}

\section{Methodology}\label{se:MethodEg}
In this section, the considered methods for obtaining the initial estimates of $\theta$ are briefly explained. Then their influence on the success rate of the maximum likelihood estimator is described.
%}

\subsection{Initial Estimates Procedure}\label{se:initEstProcs} %{Obtaining initial estimates}

%\begin{itemize}
%\item
%Obtain a good estimate of the FRF of the system by using the LPM: short description + reference

\begin{figure}
  \centering
  \includegraphics
  [width=0.95\columnwidth]{\thisDir/fig/flowgraph}
  \caption[Flow chart of different initialization procedures.]{Flow chart depicting the different estimation procedures, from left to right: (1) an approach using the true model as initial estimate; (2) an existing approach using \gls{BTLS} and \gls{GTLS} starting values, and the novel methods outlined in this chapter (3, 4).
  The flow of non-parametric data is depicted by full arrows, with the \gls{FRF} data marked by asterisks. Parametric models are indicated by dashed, unfilled arrows.} \label{fig:flowgraph}
\end{figure}

The ultimate aim is to find initial estimates that are good enough to steer clear of the local optima during the \gls{MLE} optimization process in the parametric identification of each system. 
The requisite procedures for the estimators are as follows, from left to right of the flowchart in \figref{fig:flowgraph}:

\begin{enumerate}
\item
\emph{Using the true model ($\true{G}$) as initial estimate}

The estimates from this procedure are for comparison purposes with those from the other procedures. They will be crucial to the computation of the global optimum of the \gls{MLE}.

\item
\emph{Quadratic approximations of the ML}

%\JL{The Generalized Total Least Squares (GTLS) and the Bootstrapped Total Least Squares (BTLS) have been discussed in \citep{pintGuil1998}. These are modifications of \eqref{eq:MLEcf}, which still take into account the noise information, while keeping the quadratic nature w.r.t.\ $\theta$. These estimators preserve consistency (which is an asymptotic behaviour, i.e.\ for $N\to\infty$). Their finite sample behaviour is prone to be improved by the FRF smoothing tools, described below.}

The \gls{GTLS} and the \gls{BTLS} have been discussed in \citep{Pintelon1998}. These are modifications of equation \eqref{eq:MLEcf}, which still take into account the noise information while retaining the quadratic nature w.r.t. $\theta$. 
These estimators preserve consistency, such that the estimates converge asymptotically to the true parameters for $N\to\infty$. 
Their finite sample behavior, however, is suitable for improvement by the \gls{FRF} smoothing tools as described below.

\item
\emph{Local polynomial method (LPM) with truncation}

A reliably good estimate of the FRF of the chosen system can be obtained via the LPM, which is summarized as follows~\citep{Lumori2014TIM}. Formulated as a nonparametric linear-least-squares-estimate, the LPM is first applied to estimate the FRF from a full data record of a SISO system, systematically expressed in an output-error framework, shown in \figref{fig:oesetup}. The smooth characteristics of both the exact \gls{FRF} $G_0$ and the transient term $T$ allow for an optimal application of the \gls{LPM}, leading to a smooth \gls{FRF} estimate with the transient suppressed.

  With reference to a detailed discussion in~\citep{Lumori2014TIM}, the \gls{LPM} utilizes the following quadratic local polynomials to approximate (in least squares sense) and, thus, smooth $G_0$ and $T$ around a central frequency $\Omega_{k}$ in a local frequency band $r\in\mathbb{W}_n$
\begin{subequations}\label{lpmImplQuadG}
\begin{align}
G_0(\Omega_{k+r})&\approx \hat G_k + g_{1,k} r + g_{2,k}r^2
\\
T(\Omega_{k+r})&\approx T(\Omega_k)+t_{1,k}r + t_{2,k}r^2\label{lpmImplQuadT}
\end{align}
\end{subequations}
where, in general,$\ r\in\mathbb{W}_n,\quad$$\mathbb{W}_n = \{-n,-n+1,\dots,n\}$ and $n$ is a tunable parameter. 
In this instance, $n=3$.

  The \gls{LPM} estimate of the \gls{FRF} at frequency index $k$ is the first estimated local parameter, \emph{viz}:
\begin{equation}
\hat{G}_\text{poly}(\Omega_k) = \hat{G}_k
\text{.}
\end{equation}
This procedure is repeated for all $k$ in the frequency band of interest.

%The estimation of $\hat{G}_\text{poly}(\Omega_k)$ at a single frequency $\Omega_k = 0.586$~Hz is illustrated in Fig. \ref{LPM_Schematic_EG}.

\emph{Impulse response truncation}. 
%\comment{JL: in this paper, the impulse response truncation as an initial estimator is considered together with the LPM. So I would not start a new subsection for it.}
The estimate is smoothed further by truncating its impulse response function $g_{\text{poly}}(t) = \mathrm{IDFT}\, \left( \hat{G}_\text{poly}(\Omega_k)) \right)$ as originally presented in~\citep{Lumori2014TIM}.
%\comment{JL: $\hat{G}_\text{poly}(\Omega_k))$ is undefined}.

The impulse response is truncated after the time $\tau_{\mathrm{trunc}}$ where the signal becomes indistinguishable from the noise, i.e.
\begin{equation}
  g_{\mathrm{trunc}}(t) = 
  \begin{cases}
    g_{\text{poly}}(t) & t \leq \tau_{\mathrm{trunc}} \\
    0                    & t \geq \tau_{\mathrm{trunc}}
  \end{cases}
  \text{.}
\end{equation}
To this end, an estimate of the envelope of impulse response is determined by fitting an exponential function $g_{\mathrm{exp}}(t) = B e^{\beta t}$ to the peaks of the impulse response $\abs{g_{\mathrm{poly}}(t)}$ using a linear least-squares approach.
Then, $\tau_{\mathrm{trunc}}$ is determined as the time instant where this envelope function sinks below the noise level $\sigma_g$ of the impulse response, i.e. 
\begin{equation}
  \tau_{\mathrm{trunc}} = \min \left\{  t :  g_{\mathrm{exp}}(t) < \gamma \sigma_g \right\}
  = \beta^{-1} \ln \frac{\gamma \sigma_g}{B}
  \text{,}
\end{equation}
where $\gamma=1$ was used in this chapter for simplicity.
By changing $\gamma$, the user can fine-tune the bias/variance trade-off of the estimated FRF $G_{\mathrm{trunc}}(\Omega_k) = \mathrm{DFT} \left( g_{\mathrm{trunc}}(t)\right)$ further.
This can lead to a significant improvement over the classical LPM~\citep{Lumori2014TIM}.

\item \emph{\Glsfirst{RFIR}}

The \gls{RFIR} method is a special case of the \gls{RARX} method. 
%This is a well known method which is included in this paper for comparison purposes with the smooth LPM and with other existing estimators.

%\JL{
The \gls{RFIR} estimator is formulated in the time domain. 
It estimates the impulse response of a discrete time system as the minimizer of the following regularized least squares objective function:
\begin{align}\label{eq:hRFIRdef}
\hat g_{\mathrm{RFIR}} &= \argmin_{g} \norm{y - g \ast u }^2 + \sigma^2 g^{\TT}P^{-1}g
\end{align}
where $g$ is the vectorized impulse response $g(t)$, with $t = 0,1,\cdots,n_h - 1$, assuming that the impulse response is $n_g$ samples long. 
Furthermore, $g\ast u$ is the convolution of $g$ with the input signal $u$, and $P\in \mathbb{R}^{n_g\times n_g}$ is the kernel matrix. 
Note that \eqref{eq:hRFIRdef} is quadratic in $g$, such that $\hat g_{\mathrm{RFIR}}$ can be computed analytically in a single step.
The kernel matrix $P$ embodies prior knowledge on the system to be estimated. 
Here, the \gls{DCkernel} is used. 
Specifically, the element at $t_1,t_2$ of $P$ is given by
\begin{equation}
P(\alpha,\beta)_{t_1,t_2} = e^{-\alpha|t_1 - t_2|}e^{-\beta(t_1 + t_2)/2}
\end{equation}
 for $t_1,t_2 \in \left\{0,1,\cdots,n_h - 1\right\}$ which makes the estimated impulse response to be an exponentially decaying function of $t$ with decay rate $\beta^{-1}$, and correlation length $\alpha^{-1}$. 
 The latter \emph{hyperparameters} $\alpha$ and $\beta$ are determined using Empirical Bayes, i.e. as the maximizers of the \gls{LML} of the measured output signal:
\begin{align}
\mathrm{LML}(y) = -y^{\TT}\Sigma(\alpha,\beta)^{-1}y - \log |\Sigma(\alpha,\beta)| \text{,}
\end{align}
where $\Sigma(\alpha,\beta) = \sigma^2 \I + \phi^{\TT}P(\alpha,\beta)\phi$ and $\phi$ is a Toeplitz matrix constructed with $u$. 
This problem is solved using \code{fmincon} in \MATLAB.

An implementation of the \gls{RFIR} estimator is available as the \code{arx} function in the System Identification Toolbox since \MATLAB~2013b with $\mathtt{na} = 0$, $\mathtt{nb} = n_g$, $\mathtt{nk}=0$ and regularization enabled in \code{arxoptions}. 
The hyperparameters can be determined via \code{arxregul}.
More information about the \gls{RFIR} estimator can be found in
~\citep{Pillonetto2010,Chen2012}. 
The \gls{DFT} of the obtained regularized estimate of the impulse response $\hat g_{\mathrm{RFIR}}$ of the system yields a smoothed estimate of the FRF.
%\JL{Having obtained a regularized estimate of the impulse response of the system, its DFT (Discrete Fourier Transform) yields an estimate of the FRF.}
\end{enumerate}

\begin{remark}
Note that methods 3) and 4) result in a non-parametric estimate of the transfer function, represented as the \gls{FRF}. 
A corresponding parametric estimate is obtained as follows. 
A parametric estimator (GTLS, BTLS, and MLE) is invoked, where the input spectrum is considered to be $1$ at all frequencies (Dirac in the time domain), and the output spectrum is set equal to the estimated FRF, with the transient term $I$ set to zero. It is important to note that invoking the GTLS, BTLS and MLE on the FRFs from methods 3) and 4) will yield a different result from applying the GTLS, BTLS and MLE on the raw data. This is because the FRFs from 3) and 4) have been smoothed and, thus, have a significantly lower noise variance.
\TODO{spectra als formule geven}
\end{remark}
%}

%\comment{JL: one could remove the subsection below (why? see my comments below), and mention that the the above initial estimates are then fed to the MLE, with the raw data, to obtain the final estimates.}

%\subsection{Genesis of the Initial Estimates} 
%\comment{JL: why `genesis'? This section is actually where you use the initial estimates, not their origin.}
%{Deploying the initial estimates}

%\comment{ML: Fig. \ref{fig:flowgraph} should include an additional path for RFIR/RARX, and the accompanying text (below) is to be modified accordingly.}
%Fig. \ref{fig:flowgraph} shows a flowchart for the MLE implementation, using four sets of initial values consisting of the known raw input/output data, plus: \comment{JL: is it really required to repeat the content of the flowchart? It has actually been explained in further details in \sectref{se:initEstProcs}.}
%\begin{enumerate}
%\item (Left (first) path of the flowchart): The true model $G_0$ as initial values,
%\item (Second path from the left of the flowchart): Initial values obtained from \emph{existing} initial estimate strategies, namely GTLS (generalized total least squares) and BTLS (bootstrapped total least squares),
%\item (Third path from the left of the flowchart): The LPM-truncated initial values with BTLS and GTLS, and  
%%\item (Right (fourth) path of the flowchart): The RFIR initial values with BTLS and GTLS.  
%\end{enumerate}

%\JL{
%The main contribution of this paper is to demonstrate that $\hat G_\mathrm{trunc}$ (third path) yields an improvement over $\hat G_\mathrm{exist}$ (second path). \comment{JL: these symbols haven't been introduced yet. I'm getting quite confident that this subsection can be removed.}The quantification of this improvement is formally introduced in the next section.
%}

%The latter 2 ....schematic is configured to include connections of the initial values and the input to the LPM block and/or to the MLE block(s) for estimation purposes.
%\end{itemize}
%\section{Example}\label{se:Example}

\subsection{Success Rates of the Initial Estimates}

The initial estimates described above are then fed to the MLE together with the raw data to obtain the final estimates.

Formally, let $\model{\bullet}$ be the final parametric estimate of the system, where the subscript $\bullet$ denotes the methods (from left to right in Fig.~\ref{fig:flowgraph}) when an initial estimate $\model[init]{\bullet}$ is used, \emph{viz}:
%\comment{JL:be consequent with the orders (and refer to the specific branches in the flowchart)}
\begin{itemize}
\item 
$\model{0}$ via the true model $\true{G}$ as an initial estimate,
\item
$\model{exist}$ is obtained via GTLS and BTLS,
\item
$\model{trunc}$ by the use of the LPM with truncation as an initial estimate,
\item
$\model{RFIR}$, initialized by means of the RFIR method. 
\end{itemize}

A particular initial estimate is deemed \emph{successful} if the optimization of the MLE cost with the raw input-output data inserted in equation~\eqref{eq:MLEcf}, reaches the \emph{best local optimum} when iteration is initiated with the selected initial value/estimate.

\begin{conjecture}\label{conj1}
The true model $\true{G}$ as an initial estimate is the best possible initial estimate one can use for the nonlinear identification algorithm (the leftmost path in \figref{fig:flowgraph}) and will allow the optimization to converge to the \emph{best local optimum}, i.e. $\model{0}$.
\end{conjecture}
\TODO{Johan: merk op dat dit niet zo hoeft te zijn! zelfs vanuit $\true{G}$ kan het misgaan}

From this conjecture, use of the true system as an initial estimate would engender the best final estimate or hopefully even the global optimum.
Since the true system is not really known in practice, this chapter compares the capability of different possible initial estimates to emulate the result that would be obtained by using the true model as an initial estimate.

The capability is quantified by defining the success rate of the initial estimate as the probability that the identification algorithm reaches the \emph{best local optimum} when the selected initial estimate is used. %using the selected initial estimate.

%$\hat G_\mathrm{trunc}$ and $\hat G_\mathrm{BTLS}$ are discussed later on.

The success rate $\eta$ is expressed mathematically as a probability function, consistent with Conjecture \ref{conj1}, \emph{viz}:
%\JL{in agreement with assumption \ref{assumption_G0hat_is_global_optimum}} \emph{viz}:
\begin{align}
\eta=\Prob{\hat G_\bullet=\hat G_0}
\end{align}
In practice, and since the iterative algorithm usually does not reach the local optimum precisely, the above definition is relaxed to the form
\begin{align}\label{eq:successrateTol}
\eta=\Prob{ \norm{\hat G_\bullet-\hat G_0} < tol}
\end{align}
with $tol$ a numerical tolerance that will be specified later on.
\TODO{Johan: beter een relatief criterium gebruiken (delen door $\norm{\true{G}}$)}

% \comment{JL to EG: It is better to implement the success rate as given in \eqref{eq:successrateTol}, because that can more easily be adhered. Note that it is not obvious to determine a tolerance on the difference between the estimated and the true model, because that depends on the noise variance of the estimate. On the other hand, the tolerance in \eqref{eq:successrateTol} and \eqref{eq:toldef} is independent of the noise variance.
%\comment{\EG{it is implemented that way now}}
% }

%\begin{align}
%p(\hat G_x=\hat G_\circ) %G_0(q)= B(q)/A(q)
%\end{align}

The distance measure used in equation \eqref{eq:successrateTol} is defined, for discrete time, as
\begin{align}
\|G\|_2 = \sqrt{\frac{1}{2\pi}\int_0^{2\pi} |G(e^{-j\omega})|^2\ \mathrm{d}\omega}
\end{align}
 which, in practice, is computed by using the adaptive global quadrature algorithm provided by the MATLAB function \code{integral}.
%\comment{EG: The exact implementation is in the L2Norm function. Basically, either numerical or symbolic integration can be used (as our models are parametric), but both yield equally workable results (cfr. testDistance). I picked a full numerical approach as the symbolic one doesn't always work on my computer.}
%\JL{The success rate in \eqref{eq:successrateTol} will be estimated via Monte Carlo simulations, further\ML{/later} on.}
The success rate in equation \eqref{eq:successrateTol} is estimated via Monte Carlo simulations in  Section \ref{se:CompuSR}.

% !TEX root =  InitialValuesPaper.tex

\section{Demonstration}\label{se:Demo}

\subsection{The System Under Consideration}

Two systems are considered. The first, with a quality factor $Q = 6\unit{dB}$, has the transfer function
  \begin{equation}
     G_0(z) \Big|_{Q=6\unit{dB}}
    = 1.74 \cdot 10^{-3}
    \frac{ z^2 + 2 z + 1 }
         { z^2 - 1.93 z + 0.94}
    \label{eq:systemundertest}
    \text{.}
  \end{equation}
The second system, with a quality factor $Q = 20\unit{dB}$, has the transfer function
  \begin{equation}
    G_0(z) \Big|_{Q=20\unit{dB}}
    = 309 \cdot 10^{-6}
            \frac{z^2 + 2 z + 1}
                 {z^2 - 1.98 z + 0.989}
    \label{eq:systundertest-20dB}
    \text{.}
  \end{equation}

Both systems are low-pass Chebyshev filters shown in \figref{fig:exampleFRF}. 
Their inputs are excited by zero mean white Gaussian noise with unit variance, and the outputs are disturbed by different realizations of white gaussian noise, as in \figref{fig:oesetup}, with a variance $\sigma_v^2$, such that a prescribed \gls{SNR} defined as
\begin{equation}
  \mathrm{SNR} 
    = \frac{\sigma_v}
           {\rms{y_0}}
    \approx \frac{\sigma_{v}}
                 {\left\| G_0 \right\|_2 \rms{u_0}}
    =  \frac{\sigma_{v}}
            {\left\| G_0 \right\|_2}
  \label{eq:SNR-definition}
  \text{.}
\end{equation}
is attained.
$\rms{x} = \sqrt{N^{-1} \sum_t x^2(t)}$ denotes the root-mean-squared value of $x$.
%}%}

%\JL{
%\subsection{Successful VS failed estimation}
\subsection{Comparison of Successful and Failed Estimates}

%Referring to \figref{fig:exampleFRF}, 4 FRFs are shown: $G_0$, $\hat G_0$, one successful $\hat G_\bullet$ and one failed $\hat G_\bullet$.
%}
%\EG{
 \figref{fig:exampleFRF} depicts the Bode plot of the actual system $G_0$, in black
%\comment{JL: in what color? (EG) black},
 and different estimates $\model{exist}$
 %\comment{JL: does it matter whether it is $\hat G_\mathrm{exist}$ or any other $\hat G_\bullet$? (EG) No, other methods will have similar results} 
 that may or may not approximate $\hat{G}_{0}$ well, according to the distance defined by equation \eqref{eq:successrateTol}.
 The systems shown here are drawn randomly from the successes and failures observed in the Monte Carlo simulations, discussed later.

 It is clear from \figref{fig:exampleFRF} that the successful estimates virtually coincide with the true system. This means 
that their distances from the true system, given by $\norm[2]{\model{\bullet} - \true{G}}$, have small values. 
On the other hand, the `failed' estimates are inaccurate descriptions of the true system. It is apparent that their associated distances from the true system are larger by, at least, an order of magnitude than for the successful estimates.
   %Given the 2-norm \eqref{eq:successrateTol} employed 
   %Using the 2-norm in equation \eqref{eq:successrateTol} to classify `successes' and `failures', it can be observed that the `successes' approximate $\hat{G}_0$ well w.r.t. the relative error, except when $\abs{G_0(\omega)}$ is small.
 %Practically, 
 %In practice, this means that a successful estimate provides reasonable estimates of the pass-band of the system, but may perform poorly in the stop-band. \comment{Further discussion required: what about the absolute error? Isn't it more or less white over the complete frequency band?} \ML{To be discussed}

 Estimation `failures' can be attributed to the optimization procedure getting stuck in local optima, resulting in very poor estimates of $G_0(\omega)$ over broad frequency ranges, as seen in \figref{fig:exampleFRF}. 
 %In \figref{fig:exampleFRF}, it can be seen that these \ML{`failures' result inprovide a very poor estimate of $G_0(\omega)$ over the broad frequency ranges.
 %Consequently, such estimates are practically worthless as an approximation of $G_0$.
 Consequently, such estimates are practically worthless approximations of $G_0$.
 %}

\begin{figure}
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{0.6\figurewidth}
  \input{\thisDir/fig/exampleSystems.tikz}
 % \caption{Results of an example of Monte Carlo simulation.}
 \caption[Bodeplot of successful ad failed estimates.]{Illustrative Bode plots of some successful~\legref{leg:example-good-estimate} and failed~\legref{leg:example-bad-estimate} estimates $\model{\bullet}$ for both systems $G_0$~\legref{leg:example-exact} under test.
 Success or failure is determined using the criterion in equation~\eqref{eq:successrateTol}. The $3\unit{dB}$ bandwidth of each system is highlighted~\legref{leg:example-3db-bandwidth}.}
\label{fig:exampleFRF}
\end{figure}

%\subsection{Computing the success rates}
\subsection{Computation of  the Success Rates}\label{se:CompuSR}
